Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference

### 2024-04-12 paper/paper2_neo/paper.tex
Update paper test data section

Multiple efforts have been made in this direction, treating spectra in 
different ways. Some take the wavelength and flux values and construct
neural networks to create this mapping between stellar parameters and the
observed spectra, such as [EXAMPLE] and [EXAMPLE]. Others take a more 
physics-informed approach by decomposing spectra into some eigenbasis and 
treating it as a regression problem, such as \texttt{starfish}.
However, \texttt{blase} takes a different approach. It treats spectra not as 
a set of pixels or a set of eigenbasis coefficients but as a set of 
spectral lines, specifically Voigt profiles. There is no metric of this 
approach being better or worse than the others mentioned so far, but it is 
a different paradigm and is worth exploring as a semi-empirical tool for
spectral inference. This study aims to take advantage of \texttt{blase}'s
ability to change a spectrum's basis from flux versus wavelength into
spectral line versus spectral line properties, and use that to develop a
robust tool for stellar parameter inference.

@@

\begin{table}[h]
    \hspace*{-1.1cm}\begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\textbf{PHOENIX Grid Subset}}\\
        \hline\hline
        Parameter & Symbol & Range\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0.5] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [0, 5]\\
        Wavelength & $\lambda$ & [8038, 12849] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. These limits 
    were imposed to reduce the computational cost of the algorithms. The 
    wavelength limits in particular roughly line up with that of the 
    Habitable Zone Planet Finder (HPF) spectrograph.}
\end{table}

@@

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was when pickled 
and written to a \texttt{.pkl} file. These interpolators represent multiple 
manifolds mapping fundamental stellar parameters to spectral line properties. 
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuous one. With the given size of the PHOENIX subset, the 
interpolator list takes up 19 GB of disk space. Since this is now evaluable,
we call this the PHOENIX generator.

\section{Bayesian Inference and Testing}
\subsection{Inference Algorithm}
With the PHOENIX generator complete, we used Bayesian optimization as the
inference algorithm, specifically the \texttt{gp\_minimize} function from
the \texttt{scikit-optimize} library. This algorithm uses a Gaussian Process
to model the objective function, which in this case was the $\chi^2$ loss
function between the interpolated spectrum and the true spectrum. The 
optimizer was configured to first run 10 random evaluations to seed the
surrogate model, then run 90 more evaluations now guided by the surrogate
model. This totals to 100 evaluations, which was deemed sufficient for this
study.

\subsection{Test Data}
Text

@@

\section{Discussion}
The computer used for this study, Triton, has the following specifications:
\begin{table}[h]
    \hspace*{0.7cm}\begin{tabular}{|c|c|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Triton Specs}}\\
        \hline\hline
        CPU & AMD EPYC 7513\\
        Memory & 256 GB\\
        GPU & Nvidia A100 40GB ($\times$2)\\
        \hline
    \end{tabular}
    \caption{This computer is owned by Caroline Morley and is available to
    members of her research group. It was used for all computations, but not 
    for generating visualizations.}
\end{table}

@@

    the [Fe/H] range to [-4.0, 1.0] dex, but also includes the alpha element 
    abundance, which we elected to fix at 0 for this study. In addition to 
    the actual fundamental stellar parameters, we also took a subset of the 
    PHOENIX wavelength range, with the full [500, 55000] \AA \ wavelength 
    range also being left to future work.

@@

end of the day, we aim to develop a tool that can analyze spectrum based on 
how nature behaves rather than how data behaves.

>>>>

Multiple efforts have been made in this direction, treating spectra in 
different ways. Some take the wavelength and flux values and construct
neural networks to create this mapping between stellar parameters and the
observed spectra, such as \texttt{IN-SYNC}. Others take a more 
physics-informed approach by decomposing spectra into some eigenbasis and 
treating it as a regression problem, such as \texttt{starfish}.
However, \texttt{blase} takes a different approach. It treats spectra not as 
a set of pixels or a set of eigenbasis coefficients but as a set of 
spectral lines, specifically Voigt profiles. There is no metric of this 
approach being better or worse than the others mentioned so far, but it is 
a different paradigm and is worth exploring as a semi-empirical tool for
spectral inference. This study aims to take advantage of \texttt{blase}'s
ability to change a spectrum's basis from flux versus wavelength into
spectral line versus spectral line properties, and use that to develop a
robust tool for stellar parameter inference.

@@

\begin{table}[h]
    \hspace*{-1.1cm}\begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\textbf{PHOENIX Grid Subset}}\\
        \hline\hline
        Parameter & Symbol & Range\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [2, 6]\\
        Wavelength & $\lambda$ & [8038, 12849] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. These limits 
    were imposed to reduce the computational cost of the algorithms and to 
    ensure a rectilinear parameter space in order to work with the chosen
    interpolator backend. The wavelength limits in particular roughly line 
    up with that of the Habitable Zone Planet Finder (HPF) spectrograph.}
\end{table}

@@

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was when pickled 
and written to a \texttt{.pkl} file. These interpolators represent multiple 
manifolds mapping fundamental stellar parameters to spectral line properties. 
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuous one. With the given size of the PHOENIX subset, the 
interpolator list takes up 13.2 GB of disk space. Since this is now evaluable,
we call this the PHOENIX generator.

\section{Bayesian Inference and Testing}
\subsection{Inference Algorithm}
With the PHOENIX generator complete, we used Bayesian optimization as the
inference algorithm, specifically the \texttt{gp\_minimize} function from
the \texttt{scikit-optimize} library. This algorithm uses a Gaussian Process
to model the objective function, which in this case was the RMS 
(Root-Mean-Square) loss between the interpolated spectrum and the true 
spectrum. The optimizer was configured to first run 30 random evaluations to 
seed the surrogate model, then run 20 more evaluations now guided by the 
surrogate model. This totals to 50 evaluations, which was deemed sufficient 
for this study. We then decided to use ensemble methods, running 32 parallel
optimizers to get a more statistically sound result. One inference run 
takes an average of [TIME] to complete on Triton.

\subsection{Test Data}
To test the performance of the inference algorithm, we used the PHOENIX subset
itself. At first glance, this may seem like a useless endeavor, as the 
PHOENIX generator has indeed memorized the PHOENIX subset, being able to
reconstruct a PHOENIX spectrum when evaluated at that grid point. However,
that is precisely what allows us to use the PHOENIX subset as test data.
The Bayesian optimizer's surrogate model is seeded by random continuously 
sampled generator evaluations within the search space, \textit{not} grid 
points of the PHOENIX subset, meaning the surrogate model has no `memorization' 
to speak of. If the optimizer had been an iterative grid refinement search, 
this would not have been possible, because grid methods rely on just that, 
the grid.

@@

\section{Discussion}
The computer used for this study, Triton, has the following specifications:
\begin{table}[h]
    \hspace*{0.6cm}\begin{tabular}{|c|c|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Triton Specs}}\\
        \hline\hline
        CPU & AMD EPYC 7513\\
        Memory & 256 GB\\
        GPU & Nvidia A100 40GB ($\times$2)\\
        \hline
    \end{tabular}
    \caption{This computer is owned by Caroline Morley and is available to
    members of her research group. It was used for all computations, but not 
    for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a 
    boost clock of 3.65 GHz.}
\end{table}

@@

    the [Fe/H] range to [-4.0, 1.0] dex and the $\log(g)$ range to [0, 6], 
    but also includes the alpha element abundance parameter, which we 
    elected to fix at 0 for this study. In addition to the actual fundamental 
    stellar parameters, we also took a subset of the PHOENIX wavelength range, 
    with the full [500, 55000] \AA \ wavelength range also being left to 
    future work.

@@

end of the day, we aim to develop a tool that can analyze spectra with the 
advent of physics-informed machine learning.
